Report ASP20 Boost

### 2020-06-05 Reports/report_1/1_REPORT.Rmd
added Derivatives Gamma and  Beta
minor updates to report_1

The course "Advanced Statistical Programming with R" consists in implementing a package to address location-scale regression by building upon an R6 class [@R-R6] given to us as part of the `asp20model` repository [@R-asp20model]. 
The sample dataset consists of a response $y$, i.e, the expectation is in linear relationship to a set of predictors $x_i$. 
Its variance, in contrast, is dependent on a different set of linear predictors $z_i$, however transformed by the response function $g(x) = exp(x)$. 
This ensures positiveness of the variances. 
The goal approach is to estimate the effects of both: $x_i$ on the expectation of $y$ (location) - this is captured by the estimates $\beta$, and $z_i$ on the variance of $y$ (scale) - this is captured by the estimates $\gamma$.
Our student group `asp20boost` solves this task applying the concept of boosting.

@@

The boosting concept relies on the idea of iterative estimation. 
To estimate the location parameters $\beta$ via boosting, one starts with initial estimates, which may be far from optimal. 
However, with estimates and the response vector at hand, residuals can be calculated. 
The boosting algorithm focuses on those calculated residuals.
The effect of the predictors $x_i$ on these residuals is estimated using least squares and the resulting effect is added to the location-estimate $\beta$ adjusting it  by means of the learning rate $\nu$. 
This yields a better fit of the location-model, which results in smaller residuals.
> These new residuals are estimated again in the next iteration, yielding smaller effect sizes, and hence a convergence of the boosting algorithm towards the OLS-estimate.

@@

To implement a boosting algorithm we constructed the function `gradient_boost` as a part of our R6-Class `LocationScaleRegressionBoost`. which  directly builds upon the `LocationScaleRegression` R6 class of the `asp20model`package set at our disposal for this course and this purpose specifically [@R-asp20model].

@@

A simple booster for the location parameters $\beta$ works without extending the `asp20model`'s `LocationScaleRegression` class. 
The current residuals may be extracted with the `resid()` command and then estimated.
The location parameter is updated. 
This causes the `LocationScaleRegression` class to calculate updated state-dependent objects, such as the log likelihood, the gradients and the residuals. 
This objects are again used to repeatedly  estimate residuals and update the location parameter $\beta$.

The simple booster for the scale parameters $\gamma$, as already mentioned, consists in repeating the same two steps of first, estimating a term and then updating the scale parameters $\gamma$ by adding the estimated effect - adjusted by the learning rate $\nu$. 
The estimated term here is the derivative of the log likelihood of the normal distribution.
To calculate this, our code makes use of the `resid-function()`, but this time passing the argument "deviance", which results in residuals adjusted by the fitted scale estimates.
This simple implementation of a boosting algorithm for location and scale served the code-basis or "skeleton" for our package and can, hence, be thought as the file milestone of this project.

@@

A useful extension of our simple boosting algorithm regards *componentwise boosting*. 
The key idea here is to not update a whole parameter vector, but only one entry of it.
The entry chosen for the update at this juncture is the one yielding the best improvement in terms of minimization of the loss function.
> Corrected: Check again. The loss function in our case is the deviance calculated using $\gamma$ and our $\beta$.

@@

We implement this in our package by the extending the `LocationScaleRegression` class to include the two active fields `bestFittingVariableBeta` and `bestFittingVariableGamma`.

@@

## Application

[Johannes: Munich Rent Data? - Identify Outliers]
If you want to cite Kneib to include this copy the following and update the page number:  @Fahrmeir.2013, 217-219.

>>>>

The course "Advanced Statistical Programming with R" consists in implementing a package to address location-scale regression by building upon an R6 class [@R-R6] given to us as part of the `asp20model` repository [@R-asp20model]. 
The sample dataset consists of a response $y$, i.e, the expectation is in linear relationship to a set of predictors $x_i$. 
Its variance, in contrast, is dependent on a different set of linear predictors $z_i$, however transformed by the response function $g(x) = exp(x)$. 
This ensures positiveness of the variances. 
The goal is to estimate the effects of both: $x_i$ on the expectation of $y$ (location) - this is captured by the estimates $\beta$, and $z_i$ on the variance of $y$ (scale) - this is captured by the estimates $\gamma$.
Our student group `asp20boost` solves this task applying the concept of boosting.

@@

The boosting concept relies on the idea of iterative estimation. 
To estimate the location parameters $\beta$ via boosting, one starts with initial estimates, which may be far from optimal. 
However, with estimates and the response vector at hand, residuals can be calculated. 
The boosting algorithm focuses on those calculated residuals.
The effect of the predictors $x_i$ on these residuals is estimated using least squares and the resulting effect is added to the location-estimate $\beta$ - adjusted by the learning rate $\nu$. 
This yields a better fit of the location-model, which results in smaller residuals.
These new residuals are estimated again in the next iteration, yielding smaller effect sizes, and hence a convergence of the boosting algorithm towards the OLS-estimate.

@@

To implement a boosting algorithm we constructed the function `gradient_boost` as a part of our R6-Class `LocationScaleRegressionBoost`, which  directly builds upon the `LocationScaleRegression` R6 class of the `asp20model`package set at our disposal for this course and this purpose specifically [@R-asp20model].

@@

A simple booster for the location parameters $\beta$ works without extending the `asp20model`'s `LocationScaleRegression` class. 
The current residuals may be extracted with the `resid()` command and then estimated.
The location parameter is updated. 
This causes the `LocationScaleRegression` class to calculate updated state-dependent objects, such as the log likelihood, the gradients and the residuals. 
These objects are again used to repeatedly  estimate residuals and update the location parameter $\beta$.

The simple booster for the scale parameters $\gamma$, as already mentioned, consists in repeating the same two steps of first estimating a term and then updating the scale parameters $\gamma$ by adding the estimated effect - adjusted by the learning rate $\nu$. 
The estimated term here is the derivative of the log likelihood of the normal distribution.
To calculate this, our code makes use of the `resid()`-function, but this time passing the argument "deviance", which results in residuals adjusted by the fitted scale estimates.

This simple implementation of a boosting algorithm for location and scale served the code-basis or "skeleton" for our package and can be thought of as the first milestone of this project.

@@

A useful extension of our simple boosting algorithm regards *componentwise boosting*. 
The key idea here is to not update a whole parameter vector, but only one entry of it.
The entry chosen for the update at this juncture is the one yielding the best improvement in terms of minimization of the loss function.
The loss function in our case is the deviance calculated using $\gamma$ and our $\beta$.

@@

### 1. Beta

\begin{equation} 
\begin{aligned}
\beta_{j}^{(0)} = 0, \text{ for} j = 1,..,k. \\
u_i = y - X\hat{\beta}^{(t-1)} \\
\hat{b}_j = ((\vec{x}^j)^T \space\vec{x}^j)^{-1}((\vec{x}^j)^T \vec{u} \\
\end{aligned}
\end{equation}

### 2. Gamma

\begin{equation} 
\begin{aligned}
\gamma_{j}^{(0)} = 0, \text{ for} j = 1,..,k. \\
u_i = \text{deviance} \\
\hat{b}_j = U^2Z^2e^{-2(Z\Gamma)} \\
\end{aligned}
\end{equation}

### 3. Same procedure.

Fit separate models for all covariates i.e obtain

\begin{equation} 
\hat{b}_j = ((\vec{x}^j)^T \space\vec{x}^j)^{-1}((\vec{x}^j)^T \vec{u}, \text{ for} \space j = 1,..,k. 
\end{equation}

And determine the best fitting variable via

\begin{equation} 
j^* = \arg\min_{j = 0,..,k} \sum_{i=1}^{n}(u_i-x_{ij}\hat{b}_j)^2.
\end{equation}

### 4. Update Coefficients.

\begin{equation} 
\begin{aligned}
\hat{\beta}_{j^*}^{(1)}  = \hat{\beta}_{j^*}^{(0)}+ \nu\hat{b}_{j^*} \\
\hat{\beta}_{j}^{(1)}  = \hat{\beta}_{j}^{(0)}, j\neq j^*.
\end{aligned}
\end{equation}

We implement this in our package by extending the `LocationScaleRegression` class to include the two active fields `bestFittingVariableBeta` and `bestFittingVariableGamma`.

@@
